---
title: "Case Study 1"
author: "Jordan Salsman"
date: "10/24/2019"
output: html_document
---

### Load follwing libraries

###ggplot2 - visualizations
###tidyverse
##naniar and visdat - visualizations for missing data
###mice - Imputation


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggthemes)
library(tidyverse)
library(naniar)
library(visdat)
library(GGally)
library(mice)  
library(VIM)
library(lattice)
library(ggmap)
library(treemapify)
library(stringr)
library(class)
library(e1071)
library(klaR)
library(ROCR)
library(openintro)
```

### Load data into memory from csv 

###  Load Breweries from csv
### Load Beers from csv 
###  Merge Breweries and Beers by the Brewery
###  Rename Names column after merging two datasets

```{r }
breweries_ds <- read.csv("/Users/jordansalsman/Desktop/Breweries.csv", strip.white = TRUE)
beers_ds     <- read.csv("/Users/jordansalsman/Desktop/Beers.csv", strip.white= TRUE)
bmerged_ds   <- merge(breweries_ds,beers_ds,by.x="Brew_ID",by.y="Brewery_id")
names(bmerged_ds)[2] <- "BreweryName"
names(bmerged_ds)[5] <- "BeerName"
```

### Structures of the datasets
### breweries_ds - Breweries in USA
### beers_ds    - Different beer styles with ABV and IBU
### bmerged_ds  - Merged dataset

```{r}
str(breweries_ds)
str(beers_ds)
str(bmerged_ds)
```

### Merged dataset
### First and last 10 records from merged set
```{r}
head(bmerged_ds)
tail(bmerged_ds)
```

### Summary Statistics of both datasets
### breweries_ds - Breweries in USA
### beers_ds    - Different beer styles with ABV and IBU
### bmerged_ds  - Merged dataset

```{r}
summary(breweries_ds)
summary(beers_ds)
summary(bmerged_ds)
```

### Visual Summary
### Total Breweries per state
### Bar Chart

```{r}
b_states <- breweries_ds %>% group_by(State) %>% count() %>% arrange(desc(n))
b_states$State <- factor(b_states$State, levels = b_states$State) 
ggplot(b_states, aes(x=State, y=n)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  labs(title="Ordered Bar Chart", 
       subtitle="State Vs Number of Breweries", 
       caption="Beer Analysis") + 
  xlab("State")+
  ylab("Total Breweries")+
  theme_wsj()+
  theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
  theme(plot.title    = element_text(size = rel(0.5)),
  plot.subtitle = element_text(size = rel(0.5)),
  axis.text.x   = element_text(vjust=0.6,size=10),
  axis.title    = element_text(size = rel(0.5)),
  legend.position  = "right",
  legend.direction ="vertical",
  legend.title = element_text(size = rel(0.2)))
```

### Heat Map of Breweries per state
```{r}
states <- map_data("state")
states$region <- state2abbr(states$region)
colnames(states)[colnames(states)=="region"] <- "State"
Brew_count <- breweries_ds %>% group_by(State) %>% count(State)
full <- merge(x = states, y = Brew_count, by = "State")
ggplot() + 
 geom_polygon(data=full, aes(x=long, y=lat, group = group, fill = n),colour="black")+
 coord_fixed()+ guides(fill=guide_legend(title="Number of Breweries")) + theme_classic() +     ggtitle("The Number of Breweries Across The Country") + annotate("text",x = -120, y = 31, label = "Alaska = 7") + annotate("text",x = -120, y = 28, label = "Hawaii = 4") + geom_text(label = n)
```

### Missing Values
### breweris_ds - Breweries in USA
### beers_ds    - Different beer styles with ABV and IBU

```{r}
beer_ounces <- beers_ds
beer_ounces$Ounces <- as.factor(beer_ounces$Ounces)
gg_miss_var(bmerged_ds,facet=State)+
  theme_wsj()+
  theme(axis.text.x = element_text(angle=65, vjust=0.3))+  
  theme(plot.title    = element_text(size = rel(0.5)),
  plot.subtitle = element_text(size = rel(0.5)),
  axis.text.x   = element_text(vjust=0.6,size=8),
  axis.text.y   = element_text(vjust=0.2,size=4),
  axis.title    = element_text(size = rel(0.5)),
  legend.position  = "right",
  legend.direction ="vertical",
  legend.title = element_text(size = rel(0.2)))
```


### Imputations
### 62 records doens't have IBU and SBV both - these are deleted
### Total records to process 2410-62 =  2348
### Prediction is required on 943 missing IBU
### Total records with IBU is 1405

```{r}
bmerged_final_ds <- bmerged_ds %>% filter(!is.na(ABV) | !is.na(IBU))
non_missing_ibu <- bmerged_final_ds %>% filter(!is.na(IBU))
non_missing_ibu$imputed <- "N"
missing_ibu <- bmerged_final_ds %>% filter(is.na(IBU))
missing_ibu$imputed <- "Y"
model_ibu <- non_missing_ibu %>% lm(formula=log(IBU)~log(ABV))
summary(model_ibu)
missing_ibu$IBU <- round(exp(predict(model_ibu,missing_ibu)),0)
imputed_df <- rbind(non_missing_ibu,missing_ibu)
imputed_df$imputed <- as.factor(imputed_df$imputed)
imputed_df %>% ggplot(aes(x=ABV,y=IBU,color=imputed))+
               geom_point()+ggtitle("ABV Vs IBU") + 
               theme_wsj()+
               theme(axis.text.x = element_text(angle=65, vjust=0.3))+  
               theme(plot.title    = element_text(size = rel(0.5)),
               plot.subtitle = element_text(size = rel(0.5)),
               axis.text.x   = element_text(vjust=0.6,size=8),
               axis.text.y   = element_text(vjust=0.2,size=6),
               axis.title    = element_text(size = rel(0.5)),
               legend.position  = "right",
               legend.direction ="vertical",
               legend.title = element_text(size = rel(0.5)))
```

### State Vs Median ABV

```{r}
abv_by_states <- imputed_df  %>% group_by(State) %>%  summarise(Median=as.numeric(median(ABV))) %>% arrange(desc(Median))
abv_by_states$State <- factor(abv_by_states$State, levels = abv_by_states$State) 
ggplot(abv_by_states, aes(x=State, y=Median)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  labs(title="Ordered Bar Chart", 
       subtitle="Median ABV Vs State", 
       caption="Beer Analysis") + 
  xlab("State")+
  ylab("Alcohol content")+
  theme_wsj()+
  theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
  theme(plot.title    = element_text(size = rel(0.5)),
  plot.subtitle = element_text(size = rel(0.5)),
  axis.text.x   = element_text(vjust=0.6,size=7),
  axis.title    = element_text(size = rel(0.5)),
  legend.position  = "right",
  legend.direction ="vertical",
  legend.title = element_text(size = rel(0.2)))
```

### State Vs Median Bitterness

```{r}
ibu_by_states <- imputed_df %>% group_by(State) %>%  summarise(Median=as.numeric(median(IBU))) %>% arrange(desc(Median))
ibu_by_states$State <- factor(ibu_by_states$State, levels = ibu_by_states$State) 
ggplot(ibu_by_states, aes(x=State, y=Median)) + 
  geom_bar(stat="identity", width=.8, fill="tomato3") + 
  labs(title="Ordered Bar Chart", 
       subtitle="Median IBU Vs State", 
       caption="Beer Analysis") + 
  xlab("State")+
  ylab("Bitterness")+
  theme_wsj()+
  theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
  theme(plot.title    = element_text(size = rel(0.5)),
  plot.subtitle = element_text(size = rel(0.5)),
  axis.text.x   = element_text(vjust=0.6,size=7),
  axis.title    = element_text(size = rel(0.5)),
  legend.position  = "right",
  legend.direction ="vertical",
  legend.title = element_text(size = rel(0.2)))
```

### Max ABV

```{r}
max(imputed_df$ABV)
which.max(imputed_df$ABV)
print(imputed_df[1569,])
```


### Distribution of ABV
### boxplot Max ABV
```{r}
ibu_by_states <- imputed_df %>% group_by(State) %>%  arrange(desc(ABV))
ibu_by_states$State <- factor(ibu_by_states$State, levels = unique(ibu_by_states$State)) 
str(ibu_by_states)
ibu_by_states %>% ggplot(aes(y=ABV,x=State)) +
                  geom_boxplot(aes(fill=State), alpha=0.8,show.legend = FALSE) + 
                  labs(title="Alcohol Content", 
                  caption="Beer Analysis") + 
                  xlab("State")+
                  ylab("Alcohol content")+
                  theme_wsj()+
                  theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
                  theme(plot.title    = element_text(size = rel(0.5)),
                  plot.subtitle = element_text(size = rel(0.5)),
                  axis.text.x   = element_text(vjust=0.6,size=7),
                  axis.title    = element_text(size = rel(0.5)),
                  legend.position  = "right",
                  legend.direction ="vertical",
                  legend.title = element_text(size = rel(0.2)))
```

### Max IBU

```{r}
max(imputed_df$IBU)
which.max(imputed_df$IBU)
print(imputed_df[1569,])
```

### Distribution of IBU
### boxplot Max IBU
```{r}
ibu_by_states <- imputed_df  %>% group_by(State) %>%  arrange(desc(IBU))
ibu_by_states$State <- factor(ibu_by_states$State, levels = unique(ibu_by_states$State)) 
str(ibu_by_states)
imputed_df %>% arrange(desc(IBU))
ibu_by_states %>% ggplot(aes(y=IBU,x=State)) +
                  geom_boxplot(aes(fill=State), alpha=0.8,show.legend = FALSE) + 
                  labs(title="IBU", 
                  caption="Beer Analysis") + 
                  xlab("State")+
                  ylab("IBU")+
                  theme_wsj()+
                  theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
                  theme(plot.title    = element_text(size = rel(0.5)),
                  plot.subtitle = element_text(size = rel(0.5)),
                  axis.text.x   = element_text(vjust=0.6,size=7),
                  axis.title    = element_text(size = rel(0.5)),
                  legend.position  = "right",
                  legend.direction ="vertical",
                  legend.title = element_text(size = rel(0.2)))
```


### density plot and Histogram of ABV Distribition
### Summary Statistics

```{r}
imputed_df %>% ggplot(aes(x=ABV, fill = State)) +
               geom_histogram(bins=30) + 
               labs(title="ABV Distribution" 
               ) + 
               ylab("Alcohol content")+
               theme_wsj()+
               theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
               theme(plot.title    = element_text(size = rel(0.5)),
               plot.subtitle = element_text(size = rel(0.5)),
               axis.text.x   = element_text(vjust=0.6,size=7),
               axis.title    = element_text(size = rel(0.5)),
               legend.position  = "right",
               legend.direction ="vertical",
               legend.title = element_text(size = rel(0.2)))
summary(imputed_df$ABV)
imputed_df %>% ggplot(aes(x=ABV)) +
               geom_density(aes(fill=State), alpha=0.8) + 
               labs(title="Density Plot", 
               caption="Beer Analysis") + 
               xlab("State")+
               ylab("Alcohol content")+
               theme_wsj()+
               theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
               theme(plot.title    = element_text(size = rel(0.5)),
               plot.subtitle = element_text(size = rel(0.5)),
               axis.text.x   = element_text(vjust=0.6,size=7),
               axis.title    = element_text(size = rel(0.5)),
               legend.position  = "right",
               legend.direction ="vertical",
               legend.title = element_text(size = rel(0.2)))
```

### Scatterplot and Correlation of ABV vs IBU

```{r}
imputed_df %>% ggplot(aes(x=ABV,y=IBU))+
               geom_point(color = "Red")+
               theme_wsj()+
               geom_smooth(method="lm")+
               labs(title="ABV Vs. IBU", 
               caption="Beer Analysis") + 
               theme(axis.text.x = element_text(angle=65, vjust=0.3))+  
               theme(plot.title    = element_text(size = rel(0.5)),
               plot.subtitle = element_text(size = rel(0.5)),
               axis.text.x   = element_text(vjust=0.6,size=8),
               axis.text.y   = element_text(vjust=0.2,size=6),
               axis.title    = element_text(size = rel(0.5)),
               legend.position  = "right",
               legend.direction ="vertical",
               legend.title = element_text(size = rel(0.5)))
cor(imputed_df$ABV,imputed_df$IBU)
```




### KNN with Cross-validation and Confusion Matrix

```{r}
library(e1071)
imputed_df$Style <- toupper(imputed_df$Style)
IPA <- imputed_df %>% filter(str_detect(Style, "IPA")) 
IPA$Style <- "IPA"
ALE <- imputed_df %>% filter(str_detect(Style, "IPA" , negate = TRUE)) %>% filter(str_detect(Style, "ALE"))
ALE$Style <- "Ale"
mod_df <- rbind(IPA, ALE)

set.seed(7)
iterations = 100
k = 30
Acc_holder = matrix(nrow = iterations, ncol = k)
for(j in 1:iterations)
{
smp <- floor(0.75 * nrow(mod_df))
train_ind <- sample(seq_len(nrow(mod_df)), size = smp)
imputed_train <- mod_df[train_ind, ]
imputed_test <- mod_df[-train_ind, ]
  for(i in 1:k)
  {
  classifications <- knn(imputed_train[,c(7,8)], imputed_test[,c(7,8)], imputed_train$Style,     prob = TRUE, k = i)
  CM <- confusionMatrix(table(classifications, imputed_test$Style))
    Acc_holder[j,i] = CM$overall[1]
  }
}
MeanAcc = colMeans(Acc_holder)
which.max(MeanAcc)
k <- c(1:30)
Mean_Acc_df <- data_frame(MeanAcc)
Mean_Acc_df <- cbind(k, Mean_Acc_df)
Mean_Acc_df %>% ggplot(aes(x = k, y= MeanAcc)) +
               geom_line(color = "blue", alpha = .8) +
               labs(title="Cross-Validating Different KNN Models", 
               caption="KNN Score Across K's") + 
               xlab("K's")+
               ylab("Prediction Accuracy")+
               theme_wsj()+
               theme(axis.text.x = element_text(angle=65, vjust=0.6))+  
               theme(plot.title    = element_text(size = rel(0.5)),
               plot.subtitle = element_text(size = rel(0.5)),
               axis.text.x   = element_text(vjust=0.6,size=7),
               axis.title    = element_text(size = rel(0.5)),
               legend.position  = "right",
               legend.direction ="vertical",
               legend.title = element_text(size = rel(0.2)))
classifications <- knn(imputed_train[,c(7,8)], imputed_test[,c(7,8)], imputed_train$Style,     prob = TRUE, k = 6)
CM <- confusionMatrix(table(classifications, imputed_test$Style))
CM
```


### Further exploration
### Separating Beers into known Light and Dark
```{r}
unique(imputed_df$Style)
imputed_df$Style <- toupper(imputed_df$Style)
Blonde <- imputed_df %>%filter(str_detect(Style, "BLONDE"))
Blonde$Style <- "Light"
Light <- imputed_df %>%filter(str_detect(Style, "LIGHT"))
Light$Style <- "Light"
Cream <- imputed_df %>%filter(str_detect(Style, "CREAM"))
Cream$Style <- "Light"
Shandy <- imputed_df %>%filter(str_detect(Style, "SHANDY"))
Shandy$Style <- "Light"
Porter <- imputed_df %>%filter(str_detect(Style, "PORTER"))
Porter$Style <- "Dark"
Stout <- imputed_df %>%filter(str_detect(Style, "STOUT"))
Stout$Style <- "Dark"
Dark <- imputed_df %>%filter(str_detect(Style, "DARK"))
Dark$Style <- "Dark"
Black <- imputed_df %>%filter(str_detect(Style, "BLACK"))
Black$Style <- "Dark"
color <- rbind(Blonde,Light,Porter,Stout,Dark,Black,Cream,Shandy)
```


### Using Naive Bayes to classify Light Vs. Dark Beers
### ROC Curve plot
### Confusion Matrix
```{r}
iterations = 100
AccHolder = numeric(100)
SensHolder = numeric(100)
SpecHolder = numeric(100)
for(seed in 1:iterations)
{
  set.seed(seed)
    Indices = sample(seq(1:length(color$Style)),round(.75*length(color$Style)))
    train_color = color[Indices,]
    test_color = color[-Indices,]
    model = naiveBayes(train_color[,c("ABV", "IBU", "State")],factor(train_color$Style, labels     = c("Light", "Dark")))
    CM = confusionMatrix(table(factor(test_color$Style, labels = c("Light",   "Dark")),predict(model,test_color[,c("ABV", "IBU", "State")])))
    AccHolder[seed] = CM$overall[1]
    SensHolder[seed] = CM$byClass[1]
    SpecHolder[seed] = CM$byClass[2]
}
mean(AccHolder)
mean(SensHolder)
mean(SpecHolder)
which.max(AccHolder)
which.max(SensHolder)
which.max(SpecHolder)

set.seed(38)
Indices = sample(seq(1:length(color$Style)),round(.75*length(color$Style)))
train_color = color[Indices,]
test_color = color[-Indices,]
model = naiveBayes(train_color[,c("ABV", "IBU", "State")],factor(train_color$Style, labels = c("Light", "Dark")))
CM = confusionMatrix(table(factor(test_color$Style, labels = c("Light", "Dark")),predict(model,test_color[,c("ABV", "IBU", "State")])))
CM

pred_nb <- predict(model,test_color[,c("ABV", "IBU", "State")], type = 'raw')
pred <- prediction(pred_nb[, 2], test_color$Style)
nb.prff = performance(pred, "tpr", "fpr")
plot(nb.prff)
```



